Advancements in Moltres for Time-Dependent Multiphysics Molten Salt Reactor Modeling

### 2024-12-09 presentation/turb-moltres.tex
presentation edits

\begin{frame}
  \frametitle{Spalart-Allmaras Turbulence Model in Moltres}
  \begin{block}{\textbf{Why the Spalart-Allmaras Model for Moltres?}}
    A computationally efficient one-equation model that provides reasonable estimates for turbulent
    effects in MSRs and is tractable for large problems on small computing clusters.
  \end{block}
  \begin{block}{\textbf{Subobjectives}}
    \textbf{Implementation of a Spalart-Allmaras turbulence model in Moltres}
    \begin{itemize}
      \item Implement a Spalart-Allmaras model in Moltres
      \item Verify and validate the model against reference turbulent flow solutions
    \end{itemize}
  \end{block}
\end{frame}

@@

\begin{frame}
  \frametitle{Spalart-Allmaras Turbulence Model in Moltres}
  \begin{block}{\textbf{Model Implementation Details}}
  \begin{itemize}
    \item Built on top of an existing incompressible Navier-Stokes model in MOOSE
    \item Uses \gls{SUPG} stabilization scheme for numerical stability
    \item $f_{t2}$ turbulence trip term is togglable (Negligible difference in fully turbulent flows)
    \item Rotation correction scheme \cite{aupoix_extensions_2003, dacles-mariani_numericalexperimental_1995}
      reduces eddy viscosity in regions of rotational but non-turbulent flow
  \end{itemize}
  \end{block}
  \pause
  \begin{block}{\textbf{Verification \& Validation Tests for the Turbulence Model}}
    \begin{enumerate}
      \item Turbulent Channel Flow Verification Test at Re $\approx 13750$
      \begin{itemize}
        \item Reference results: \gls{DNS} results by Moser et al.\ \cite{moser_direct_1999}
      \end{itemize}
      \item Turbulent Pipe Flow Validation Test at Re $\approx 40000$
      \begin{itemize}
        \item Reference results: Experimental results by Laufer \cite{laufer_structure_1954}
      \end{itemize}
      \item Backward-Facing Step Flow Validation Test
      \begin{itemize}
        \item Reference results: Experimental results by Driver \& Seegmiller \cite{driver_features_1985}
        \item Reference results: Spalart-Allmaras model results from CFL3D CFD code \cite{krist_cfl3d_1998}
      \end{itemize}
    \end{enumerate}
  \end{block}
\end{frame}

\begin{frame}
  \frametitle{Spalart-Allmaras Turbulence Model in Moltres}
  \textbf{Turbulent Channel Flow Verification Test Results}
  \begin{figure}[htb!]
    \centering
    \begin{subfigure}[b]{0.32\columnwidth}
      \centering
      \includegraphics[width=\columnwidth]{channel_vel}
      \caption{Normalized velocity distribution across the channel.}
      \label{fig:channel-vel}
    \end{subfigure}
    \hfill
    \begin{subfigure}[b]{0.32\columnwidth}
      \centering
      \includegraphics[width=\columnwidth]{channel_nondim}
      \caption{Dimensionless velocity vs.\ dimensionless wall distance.}
      \label{fig:channel-nondim}
    \end{subfigure}
    \begin{subfigure}[b]{0.32\columnwidth}
      \centering
      \includegraphics[width=\columnwidth]{channel_stress}
      \caption{Normalized stress distribution across the channel.}
      \label{fig:channel-stress}
    \end{subfigure}
    \caption{Comparison of turbulent channel flow results at Re$_\tau\approx395$ against reference
    \gls{DNS} results \cite{moser_direct_1999}. The Moltres Spalart-Allmaras model agrees consistently with
    the reference data.}
    \label{fig:channel-verification}
  \end{figure}
\end{frame}

\begin{frame}
  \frametitle{Spalart-Allmaras Turbulence Model in Moltres}
  \textbf{Turbulent Pipe Flow Validation Test Results}
  \begin{figure}[htb]
    \centering
    \begin{subfigure}[b]{0.32\columnwidth}
      \centering
      \includegraphics[width=\columnwidth]{pipe_vel}
      \caption{Normalized velocity distribution across the pipe.}
      \label{fig:pipe-vel}
    \end{subfigure}
    \hfill
    \begin{subfigure}[b]{0.32\columnwidth}
      \centering
      \includegraphics[width=\columnwidth]{pipe_nondim}
      \caption{Dimensionless velocity vs.\ dimensionless wall distance}
      \label{fig:pipe-nondim}
    \end{subfigure}
    \begin{subfigure}[b]{0.32\columnwidth}
      \centering
      \includegraphics[width=\columnwidth]{pipe_stress}
      \caption{Normalized turbulent shear stress across the pipe.}
      \label{fig:pipe-stress}
    \end{subfigure}
    \caption{Comparison of turbulent pipe flow results at Re $\approx 40000$ against reference
    experimental data \cite{laufer_structure_1954}.}
    \label{fig:pipe-verification}
  \end{figure}
\end{frame}

\begin{frame}
  \frametitle{Spalart-Allmaras Turbulence Model in Moltres}
  \textbf{Backward-Facing Step Flow Validation Test Results}
  \begin{columns}
    \column{5.5cm}
    \begin{figure}[h]
      \centering
      \includegraphics[width=\columnwidth]{backstep-geom-pres}
      \caption{Backward step geometry for the turbulent BFS flow verification test. The red box indicates
      the region shown by the close-up view in Figure \ref{fig:bfs-mesh}.}
      \label{fig:backstep-geom}
    \end{figure}
    \column{5.5cm}
    \begin{figure}[h]
      \centering
      \includegraphics[width=\columnwidth]{bfs_mesh}
      \caption{Close-up view of the mesh for the BFS flow test. The step is situated at $x=110$ m
      with a height of $H=1$ m.}
      \label{fig:bfs-mesh}
    \end{figure}
  \end{columns}
\end{frame}

\begin{frame}
  \frametitle{Spalart-Allmaras Turbulence Model in Moltres}
  \textbf{Backward-Facing Step Flow Validation Test Results}
  \begin{columns}
    \column{5.5cm}
    \begin{figure}[htb!]
      \centering
      \includegraphics[width=\columnwidth]{bfs}
      \caption{Velocity magnitude distribution and streamlines around the backward-facing step. The
      streamlines illustrate the primary and secondary recirculation zones.}
      \label{fig:bfs}
    \end{figure}
    \column{5.5cm}
    \begin{table}[htb]
      \centering
      \scriptsize
      \caption{BFS flow reattachment length estimates normalized by step height $H$.}
      \begin{tabular}{l S[table-format=1.2(2)]}
        \toprule
        Source & {Reattachment length [-]} \\
        \midrule
        Experimental data & 6.26(10) \\
        CFL3D & 6.1 \\
        Moltres & 6.36 \\
        \bottomrule
      \end{tabular}
      \label{table:bfs-reattach}
    \end{table}
  \end{columns}
\end{frame}

\begin{frame}
  \frametitle{Spalart-Allmaras Turbulence Model in Moltres}
  \textbf{Backward-Facing Step Flow Validation Test Results}
  \begin{figure}[htb]
    \centering
    \hfill
    \begin{subfigure}[b]{0.24\columnwidth}
      \centering
      \includegraphics[width=\columnwidth]{bfs_downstream_vel}
      \caption{Normalized velocity distributions downstream of step.}
      \label{fig:bfs-downstream}
    \end{subfigure} %\hfill \\
%    \centering
    \hfill
    \begin{subfigure}[b]{0.24\columnwidth}
      \centering
      \includegraphics[width=\columnwidth]{bfs_cf}
      \caption{Skin friction coefficient along the bottom wall.}
      \label{fig:bfs-cf}
    \end{subfigure}
    \hfill
    \begin{subfigure}[b]{0.24\columnwidth}
      \centering
      \includegraphics[width=\columnwidth]{bfs_cp}
      \caption{Skin pressure coefficient along the bottom wall.}
      \label{fig:bfs-cp}
    \end{subfigure}% \hfill \\
%    \centering
    \begin{subfigure}[b]{0.24\columnwidth}
      \centering
      \includegraphics[width=\columnwidth]{bfs_stress}
      \caption{Normalized turbulent shear stress downstream
      of step.}
      \label{fig:bfs-stress}
    \end{subfigure}
    \caption{Comparison of backward facing step flow results against reference
    experimental data and computational data from CFL3D. $x/H$ values are normalized horizontal
    distances relative to the step.}
    \label{fig:bfs-plots}
  \end{figure}
\end{frame}

\begin{frame}
  \frametitle{Spalart-Allmaras Turbulence Model in Moltres}
  \textbf{Spalart-Allmaras Model V\&V Test Summary}
  \begin{itemize}
    \item Good agreement between Moltres and reference solutions in the turbulent channel and pipe
      flow tests
    \item Good agreement between Moltres and CFL3D in the backward-facing step flow test
    \item Some discrepancy between Moltres and experimental data in the backward-facing step flow
      test, but Moltres performed better than CFL3D at predicting flow reattachment length and
      velocity/stress distributions along $x/H=1$.
  \end{itemize}
\end{frame}

>>>>

\begin{frame}
  \frametitle{Spalart-Allmaras Turbulence Model in Moltres}
  \begin{block}{\textbf{Why the Spalart-Allmaras Model \cite{spalart_one-equation_1994} for Moltres?}}
    A computationally efficient one-equation model that provides reasonable estimates for wall-bounded
    turbulent flows and is tractable for large problems on small computing clusters.
  \end{block}
  \begin{block}{\textbf{Subobjectives}}
    \textbf{Implementation of a Spalart-Allmaras turbulence model in Moltres}
    \begin{itemize}
      \item Implement a Spalart-Allmaras model in Moltres
      \item Verify and validate the model against reference turbulent flow solutions
    \end{itemize}
  \end{block}
\end{frame}

@@

\begin{frame}
  \frametitle{Spalart-Allmaras Turbulence Model in Moltres}
  \begin{block}{\textbf{Model Implementation Details}}
  \begin{itemize}
    \item Built on top of an existing incompressible Navier-Stokes model in MOOSE
    \item Uses Streamline-Upwind/Petrov-Galerkin (SUPG) stabilization scheme for numerical stability
    \item Rotation correction scheme \cite{aupoix_extensions_2003, dacles-mariani_numericalexperimental_1995}
      reduces eddy viscosity in regions of rotational but non-turbulent flow
  \end{itemize}
  \end{block}
  \pause
  \begin{block}{\textbf{Verification \& Validation Tests for the Turbulence Model}}
    \begin{enumerate}
      \item Turbulent Channel Flow Verification Test at Re $\approx 13750$
      \begin{itemize}
        \item Reference results: \gls{DNS} results by Moser et al.\ \cite{moser_direct_1999}
      \end{itemize}
      \item Turbulent Pipe Flow Validation Test at Re $\approx 40000$
      \begin{itemize}
        \item Reference results: Experimental results by Laufer \cite{laufer_structure_1954}
      \end{itemize}
      \item Backward-Facing Step Flow Validation Test
      \begin{itemize}
        \item Reference results: Experimental results by Driver \& Seegmiller \cite{driver_features_1985}
        \& Spalart-Allmaras model results from CFL3D CFD code \cite{krist_cfl3d_1998}
      \end{itemize}
    \end{enumerate}
  \end{block}
\end{frame}

\begin{frame}
  \frametitle{Turbulent Channel Flow Verification Test Results}
  \begin{figure}[htb!]
    \centering
    \begin{subfigure}[b]{0.32\columnwidth}
      \centering
      \includegraphics[width=\columnwidth]{channel_vel}
      \caption{Normalized velocity distribution across the channel.}
      \label{fig:channel-vel}
    \end{subfigure}
    \hfill
    \begin{subfigure}[b]{0.32\columnwidth}
      \centering
      \includegraphics[width=\columnwidth]{channel_nondim}
      \caption{Dimensionless velocity vs.\ dimensionless wall distance.}
      \label{fig:channel-nondim}
    \end{subfigure}
    \begin{subfigure}[b]{0.32\columnwidth}
      \centering
      \includegraphics[width=\columnwidth]{channel_stress}
      \caption{Normalized stress distribution across the channel.}
      \label{fig:channel-stress}
    \end{subfigure}
    \caption{Comparison of turbulent channel flow results at Re$_\tau\approx395$ against reference
    \gls{DNS} results \cite{moser_direct_1999}.}
    \label{fig:channel-verification}
  \end{figure}
  \vspace{.1cm}

  $\Rightarrow$ The Moltres Spalart-Allmaras model results are consistent with reference DNS channel flow data.
\end{frame}

\begin{frame}
  \frametitle{Turbulent Pipe Flow Validation Test Results}
  \begin{figure}[htb]
    \centering
    \begin{subfigure}[b]{0.32\columnwidth}
      \centering
      \includegraphics[width=\columnwidth]{pipe_vel}
      \caption{Normalized velocity distribution across the pipe.}
      \label{fig:pipe-vel}
    \end{subfigure}
    \hfill
    \begin{subfigure}[b]{0.32\columnwidth}
      \centering
      \includegraphics[width=\columnwidth]{pipe_nondim}
      \caption{Dimensionless velocity vs.\ dimensionless wall distance}
      \label{fig:pipe-nondim}
    \end{subfigure}
    \begin{subfigure}[b]{0.32\columnwidth}
      \centering
      \includegraphics[width=\columnwidth]{pipe_stress}
      \caption{Normalized turbulent shear stress across the pipe.}
      \label{fig:pipe-stress}
    \end{subfigure}
    \caption{Comparison of turbulent pipe flow results at Re $\approx 40000$ against reference
    experimental data \cite{laufer_structure_1954}.}
    \label{fig:pipe-verification}
  \end{figure}
  \vspace{.1cm}

  $\Rightarrow$ The Moltres Spalart-Allmaras model results are consistent with reference experimental pipe flow data.
\end{frame}

\begin{frame}
  \frametitle{Backward-Facing Step Flow Validation Test Setup}
  \begin{columns}
    \column{5.5cm}
    \begin{figure}[h]
      \centering
      \includegraphics[width=\columnwidth]{backstep-geom-pres}
      \caption{Backward step geometry for the turbulent BFS flow verification test. The red box indicates
      the region shown by the close-up view in Figure \ref{fig:bfs-mesh}.}
      \label{fig:backstep-geom}
    \end{figure}
    \column{5.5cm}
    \begin{figure}[h]
      \centering
      \includegraphics[width=\columnwidth]{bfs_mesh}
      \caption{Close-up view of the mesh for the BFS flow test. The step is situated at $x=110$ m
      with a height of $H=1$ m.}
      \label{fig:bfs-mesh}
    \end{figure}
  \end{columns}
\end{frame}

\begin{frame}
  \frametitle{Backward-Facing Step Flow Validation Test Results}
  \begin{columns}
    \column{5.5cm}
    \begin{figure}[htb!]
      \centering
      \includegraphics[width=\columnwidth]{bfs}
      \caption{Velocity magnitude distribution and streamlines around the backward-facing step. The
      streamlines illustrate the primary and secondary recirculation zones.}
      \label{fig:bfs}
    \end{figure}
    \column{5.5cm}
    \begin{table}[htb]
      \centering
      \scriptsize
      \caption{BFS flow reattachment length estimates normalized by step height $H$.}
      \begin{tabular}{l S[table-format=1.2(2)]}
        \toprule
        Source & {Reattachment length [-]} \\
        \midrule
        Experimental data & 6.26(10) \\
        CFL3D & 6.1 \\
        Moltres & 6.36 \\
        \bottomrule
      \end{tabular}
      \label{table:bfs-reattach}
    \end{table}
    \vspace{.1cm}

    $\Rightarrow$ Moltres is consistent with experimental data (within uncertainty range) while CFL3D
    falls outside the uncertainty range.
  \end{columns}
\end{frame}

\begin{frame}
  \frametitle{Backward-Facing Step Flow Validation Test Results}
  \begin{columns}
    \column{11.5cm}
    \begin{figure}[htb]
      \centering
      \begin{subfigure}[t]{0.32\columnwidth}
        \centering
        \includegraphics[width=\columnwidth]{bfs_downstream_vel}
        \caption{Normalized velocity distributions downstream of step.}
        \label{fig:bfs-downstream}
      \end{subfigure} %\hfill \\
      \begin{subfigure}[t]{0.32\columnwidth}
        \centering
        \includegraphics[width=\columnwidth]{bfs_cf}
        \caption{Skin friction coefficient along the bottom wall.}
        \label{fig:bfs-cf}
      \end{subfigure}
      \begin{subfigure}[t]{0.32\columnwidth}
        \centering
        \includegraphics[width=\columnwidth]{bfs_stress}
        \caption{Normalized turbulent shear stress downstream
        of step.}
        \label{fig:bfs-stress}
      \end{subfigure}
      \caption{Comparison of backward facing step flow results against reference
      experimental data and computational data from CFL3D. $x/H$ values are normalized horizontal
      distances relative to the step.}
      \label{fig:bfs-plots}
    \end{figure}
  \end{columns}
  \begin{itemize}
    \item Minor discrepancies between numerical and experimental data typical of RANS-based turbulence models.
    \item Moltres provides more accurate measurements along $x/H=1$ than CFL3D.
  \end{itemize}
\end{frame}

\begin{frame}
  \frametitle{Spalart-Allmaras Turbulence Model in Moltres}
  \begin{block}{\textbf{Spalart-Allmaras Model V\&V Test Summary}}
    \begin{itemize}
      \item Good agreement between Moltres and reference solutions in the turbulent channel and pipe
        flow tests.
      \item Good agreement between Moltres and CFL3D in the backward-facing step flow test.
      \item Some discrepancy between Moltres and experimental data in the backward-facing step flow
        test, but Moltres performed better than CFL3D at predicting flow reattachment length and
        velocity/stress distributions along $x/H=1$.
    \end{itemize}
  \end{block}
\end{frame}
